---
title: "Mid-Semester Report"
output: pdf_document
header-includes:
    - \usepackage[labelformat=empty]{caption}
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
opts_chunk$set(echo = F)

source(here::here("R50.R"))
source(here::here("R50_power.R"))
```



## Develop a model for estimating R~50~.

I was able to use a logistic model to estimate R~50~. Below is a plot of the model and the confidence intervals.

```{r}
R50_plot
```


```{r}
R50_dat %>%
  mutate(turbines = c("None", "Present")) %>%
  kable("latex",
    caption = "$R_{50}$ Estimaiton and Confidence Intervals",
    col.names = c("Turbines", "$R_{50}$", "Lower", "Upper"),
    booktabs = T,
    escape = F,
    digits = 2
  ) %>%
  kable_styling(latex_options = c("hold_position"))
```

\newpage

## Identify impact of target conditions on R~50~.

Since the `turbines` term is significant and it represents a horizontal shift in the two curves, the estimated R~50~ will be significantly different under the two conditions.

```{r}

R50_mod %>%
  tidy() %>%
  mutate(
    p.value =
      p.value %>%
        format(digits = 1, scientific = F) %>%
        cell_spec("latex",
          background = if_else(term == "turbines",
            "yellow", "white"
          )
        )
  ) %>%
  kable("latex",
    caption = "$R_{50}$ Model Terms",
    booktabs = T,
    escape = F,
    digits = c(2, 2, 2, 2, 2)
  ) %>%
  kable_styling(
    latex_options = c("HOLD_position")
  )
```

## Develop a means of estimating power and confidence.

```{r}
power_plot
```

